---
title: "Applied Data Mining Homework 04"
author: "Xun Zhao, xz2827"
output: pdf_document
fontsize: 12pt
---

```{r setup, include = FALSE}
library(diagram)
```

# Problem 1:

The function $f$ can be interpreted as:

$$\begin{aligned}
f(x,\,y) &= I\{g(x,\,y) > 0\}\\\
g(x,\,y) &= \sigma(x) - \sigma(y)\\\
\end{aligned}$$

So the network is:

```{r, echo = FALSE, out.width = '25em', fig.align = 'center'}
names = c(expression(x), expression(y), expression(symbol(f)[1]), expression(symbol(f)[2]), expression(symbol(f)[3]))
M = matrix(nrow = 5, ncol = 5, byrow = TRUE, data = 0)
M[3, 1] = 1
M[4, 2] = 1
M[5, 3] = 1
M[5, 4] = -1
plotmat(M, pos = c(2, 2, 1), 
		curve = 0, 
		name = names, 
		lwd = 1, 
		box.size = 0.05, 
		box.type = 'circle', 
		box.lwd = 2, 
		box.prop = 1,
		shadow.size = 0, 
		arr.type = 'simple', 
		arr.lwd = 0.5, 
		dtext = 1,
		cex.txt = 0.8,
		latex = TRUE
		) 
```

Here, the three $\phi$ nodes are:

$$\begin{aligned}
\phi_1(\cdot) &= \sigma(\cdot) \\\
\phi_2(\cdot) &= \sigma(\cdot) \\\
\phi_3(\cdot) &= I\{\cdot > 0\} \\\
\end{aligned}$$

# Problem 2:

According to the definition of sigmoid function, $\sigma(x) = \frac{1}{1 + e^{-x}}$, we can get:

$$\begin{aligned}
-\sigma(-x) &= -\frac{1}{1 + e^x}\\\
&= \frac{e^x}{1 + e^x} - 1\\\
&= \frac{1}{e^{-x} + 1} - 1\\\
&= \sigma(x) - 1
\end{aligned}$$

So, the network is:

```{r, echo = FALSE, out.width = '25em', fig.align = 'center'}
names = c(expression(x), expression(symbol(f)[1]), expression(symbol(f)[2]), expression(symbol(f)[3]))
M = matrix(nrow = 4, ncol = 4, byrow = TRUE, data = 0)
M[2, 1] = 1
M[4, 2] = 1
M[4, 3] = -1
plotmat(M, pos = c(1, 2, 1), 
		curve = 0, 
		name = names, 
		lwd = 1, 
		box.size = 0.05, 
		box.type = 'circle', 
		box.lwd = 2, 
		box.prop = 1,
		shadow.size = 0, 
		arr.type = 'simple', 
		arr.lwd = 0.5, 
		dtext = 1,
		cex.txt = 0.8,
		latex = TRUE
		) 
```

Here, the three $\phi$ nodes are:

$$\begin{aligned}
\phi_1(\cdot) &= \sigma(\cdot) \\\
\phi_2(\cdot) &= c = 1 \\\
\phi_3(\cdot) &= \cdot \\\
\end{aligned}$$

# Problem 3:

According basic logical relationship:

$$\begin{aligned}
x\oplus y &= (x = y = 1) \vee (x = y = 0)\\\
&= (x \wedge y) \vee (\neg x \wedge \neg y)
\end{aligned}$$

It indicates that "`XOR`" is the "`or`" of two "`and`".

The single unit of "`and`" can be presented as $f(x,\,y) = I\{x + y > 1\}$. The single unit of "`or`" can be presented as $f(x,\,y) = I\{x + y > 0\}$. The single unit of "`not`" can be presented as $f(x) = 1 - x$.

Therefore,	

$$
\begin{aligned}
x\oplus y &= \text{or}(\text{and}(x,\,y),\,\text{and}(\text{not}(x),\,\text{not}(y)))\\\
\text{and}(x,\,y) &= I\{x + y - 1 > 0\}\\\
\text{or}(x,\,y) &= I\{x + y > 0\}\\\
\text{not}(x) &= 1 - x\\\
\text{and}(\text{not}(x),\,\text{not}(y)) &= I\{- x - y + 1 > 0\}
\end{aligned}
$$

So the network is:

```{r, echo = FALSE, fig.height = 10, fig.width = 7, out.width = '25em', fig.align = 'center'}
names = c(expression(x[1]), expression(x[2]), expression(c[1]), 
		  expression(symbol(f)[1]), expression(symbol(f)[2]), expression(symbol(f)[3]))
M = matrix(nrow = 6, ncol = 6, byrow = TRUE, data = 0)
M[4, 1] = M[4, 2] = M[5, 3] = M[6, 4] = M[6, 5] = 1
M[5, 1] = M[5, 2] = M[4, 3] = -1
plotmat(M, pos = c(3, 2, 1),
		curve = 0,
		name = names,
		lwd = 1,
		box.size = 0.05,
		box.type = 'circle',
		box.lwd = 2,
		box.prop = 1,
		shadow.size = 0,
		arr.type = 'simple',
		arr.lwd = 0.5,
		arr.pos = 0.8,
		dtext = 0.4,
		cex.txt = 0.8
		)
```

Here, the bias nodes are:

$$\begin{aligned}
c_1 &= 1\\\
\end{aligned}$$

And other three $\phi$ nodes are:

$$\begin{aligned}
\phi_1(\cdot) &= I\{\cdot > 0\}\\\
\phi_2(\cdot) &= I\{\cdot > 0\}\\\
\phi_3(\cdot) &= I\{\cdot > 0\}\\\
\end{aligned}$$

# Problem 4:

Classification function is:

$$\begin{aligned}
f(\vec{x}) &= \phi(\vec{x} \cdot \vec{w} + w_c \cdot c)\\\
&= I\{\vec{x} \cdot \vec{w} \geq c \}
\end{aligned}$$

So, the classification result of $\vec{x}$ is:

$$\begin{aligned}
f(\vec{x}) &= I\{\begin{pmatrix} -3 \\ 0 \\ \end{pmatrix} \cdot \begin{pmatrix} \frac{1}{\sqrt{2}} \\ \frac{1}{\sqrt{2}} \\ \end{pmatrix} - \frac{1}{2\sqrt{2}}>0\}\\\
&= I\{-\frac{3}{\sqrt{2}} - \frac{1}{2\sqrt{2}}>0\}\\\
&= I\{-\frac{7}{2\sqrt{2}} > 0\}\\\
&= 0
\end{aligned}$$

And the classification reuslt of $\vec{x'}$ is:

$$\begin{aligned}
f(\vec{x'}) &= I\{\begin{pmatrix} \frac{1}{2} \\ \frac{1}{2} \\ \end{pmatrix} \cdot \begin{pmatrix} \frac{1}{\sqrt{2}} \\ \frac{1}{\sqrt{2}} \\ \end{pmatrix} - \frac{1}{2\sqrt{2}}>0\}\\\
&= I\{\frac{1}{\sqrt{2}} - \frac{1}{2\sqrt{2}}>0\}\\\
&= I\{\frac{1}{2\sqrt{2}} > 0\}\\\
&= 1
\end{aligned}$$

# Problem 5:

$$\vec{x} \cdot \vec{y} = e^{\log(x_1) + \log(y_1)} + e^{\log(x_2) + \log(y_2)}$$

The network is:

```{r, echo = FALSE, fig.height = 10, fig.width = 7, out.width = '25em', fig.align = 'center'}
names = c(
	expression(x[1]),
	expression(x[2]),
	expression(y[1]),
	expression(y[2]),
	rep(expression(log()), 4),
	rep(expression(exp()), 2),
	expression(output)
	)
M = matrix(nrow = 11, ncol = 11, byrow = TRUE, data = 0)
M[5, 1] = M[6, 2] = M[7, 3] = M[8, 4] = M[9, 5] = M[9, 7] = M[10, 6] = M[10, 8] = M[11, 9] = M[11, 10] = 1
plotmat(M, pos = c(4, 4, 2, 1),
		curve = 0,
		name = names,
		lwd = 1,
		box.size = 0.05,
		box.type = 'circle',
		box.lwd = 2,
		box.prop = 1,
		shadow.size = 0,
		arr.type = 'simple',
		arr.lwd = 0.5,
		arr.pos = 0.5,
		dtext = 0.8,
		cex.txt = 0.8,
		latex = TRUE
		)
```